DATA 621 - Business Analytics and Data Mining

### 2020-12-04 Final_project/Final Project Paper - Group 2.Rmd
Big changes. Move single-date extraction into data exploration section, move creation of "exceed mean" feature to model building, put dummyVar creation AFTER train/test split in case someone wants to use exceedMean in modeling. I think it's flawed. You're using views to predict views.

# Methodology
Using the famous data set from Kaggle [@Mitchell2019TYV], we will explore
relationships between a video’s views and the number of likes, dislikes, and
comments. We may also use a video’s country of origin and the category or type
of video as predictors.

@@

# Experimentation & Results
## Data Exploration

```{r dataPrep}
# Substitute category for numeric ID through joining
us_set <- us_cats[us_set, on = 'id == category_id']

# Change data types for convenience, add target bucket, count tags and then get
# rid of actual tags and category IDs

us_mean <- mean(us_set$views)

us_set[, `:=`(trending_date = as.IDate(trending_date, format = "%y.%d.%m"),
              views = as.double(views),
              tag_count = 1L + str_count(us_set$tags, '[|]'),
              view_target = factor(ifelse(views > us_mean, 1L, 0L),
                                   levels = c(0, 1), labels = c(0, 1)),
              id = NULL,
              category = as.factor(category))
       ][, `:=`(tags = NULL,
                category = relevel(category, 'Entertainment'))]

# Make name shorter for display purposes
setnames(us_set, c('comment_count', 'tag_count'), c('comments', 'tags'))

# Categorize the variable names
boolVars <- c('comments_disabled', 'ratings_disabled', 'video_error_or_removed')
numVars <- c('views', 'likes', 'dislikes', 'comments', 'tags')
facVars <- c('category', 'view_target')
dateVars <- c('trending_date', 'publish_time')
charVars <- c('video_id', 'title', 'channel_title')

# Melt for numerics
usSetN <- melt(us_set, measure.vars = numVars, variable.factor = FALSE,
               variable.name = 'measure', value.name = 'value')

# Melt for factors
usSetF <- melt(us_set,
               measure.vars = c('category', 'view_target'),
               variable.factor = FALSE, variable.name = 'measure',
               value.name = 'value')

@@

### Factor Predictors
While tabular representation of factor predictors is difficult, a distribution
of factors by value may prove informative. It is clear from the graphs below
that interests vary by category. `Entertainment` seems to be the most common in
the US with `Music` coming in second.

```{r factorPlot1, fig.height=4.5, fig.width=8}
ggplot(usSetF[measure == 'category'], aes(y = value)) +
  geom_bar() +
  ggtitle("Distribution by Category") +
  scale_x_continuous(labels = comma)
```

```{r factorPlot2, fig.height=3, fig.width=8}
ggplot(usSetF[measure != 'category'], aes(x = value, fill = measure)) +
  geom_bar(position = 'dodge2') +
  ggtitle("Country and Global Mean Views Exceedence") +
  scale_y_continuous(labels = comma)
```

@@

## Feature Engineering & Selection
Since we are not factoring in time, it is incorrect to use all the observations.
Therefore, we will extract the latest observation by video by country and use
this subset.

```{r subsetExtraction}
# Select the row numbers of the first entry of the latest trending date by title
# and by country. There are 34 duplicates. Use this as our restricted data set
# corresponding to the most recent view count. The inner set of brackets gets
# the row number (called V1) and the outer set is a simple subset by those row
# numbers.
usExtract <- us_set[
  us_set[, .I[which.max(views)], by = c('title')]$V1
  ]
```

We will consider the relationship between `views` and the numerical predictors
of `likes`, `dislikes`, `comments`, and `tags`. We will also consider the
`category`, whether or not comments or ratings were disabled and, if there was
an error with the video.

A feature that we will also consider is `lkratio`: the ratio of likes to
dislikes. This will help ascertain if it is the magnitude or the preponderance
which has a greater effect. As there may be some rare videos with no dislikes,
the ratios will be set to 1. The variable `unan`, short for unanimous, will
reflect that some videos have likes but no dislikes, to balance the ratio of 1
which should have been `NaN`, a division by 0 error.

```{r addFeatures}
usExtract[, `:=`(lkratio = ifelse(dislikes == 0, 1, likes / dislikes),
                 unan = ifelse(dislikes == 0 & likes > 0, TRUE, FALSE))]

modDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
                   comments_disabled + ratings_disabled +
                   video_error_or_removed + lkratio + unan,
                   data = usExtract, fullRank = TRUE)
modMat <- predict(modDum, usExtract)
```

@@

```{r holdOut}
set.seed(617)
seenIDX <- createDataPartition(usExtract$views, p = 0.8)$Resample1
seenX <- modMat[seenIDX, ]
seenY <- usExtract$views[seenIDX]
hideX <- as.data.frame(modMat[-seenIDX, ])
hideY <- usExtract$views[-seenIDX]
```

### Simple Linear Regression
Linear regression may be the best known algorithm used when analyzing a
continuous numeric outcome. It searches for a linear relationship of the
predictors that minimizes the squared error between the "predictor" function and
the observations [@Sheather2009AMA]. This model will engage in stepwise feature
selection using the AIC as the optimization metric.

@@

### ElasticNet: Penalized Regression
Instead of using AIC to select features, one can make use of penalized
regression. Using an \(L_1\) penalty is the underpinnings of Lasso regression,
which can perform feature selection. Using a squared error term, \(L_2\), is at
the heart of ridge regression [@HTF2009TEo]. Using both methods together is
called the elastic net [@ZH2005RaV]. To tune the hyperparameters, which includes
the weighting between Lasso and Ridge, we will use 10-fold cross-validation.

```{r lm3Train}
trC <- trainControl(method = 'cv', number = 10L)
tG <- expand.grid(fraction = seq(0.8, 1, 0.001),
                  lambda = seq(0, 0.02, 0.001))
set.seed(181)
lm3 <- train(x = seenX, y = seenY,
             method = 'enet', trControl = trC, tuneGrid = tG)
```

@@

```{r modTest}
lm1P <- predict(lm1, newdata = hideX)
lm2P <- predict(lm2, newdata = hideX)
lm3P <- predict(lm3, newdata = hideX)
compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'ElasticNet'),
                        RMSE = c(RMSE(lm1P, hideY), RMSE(lm2P, hideY),
                                 RMSE(lm3P, hideY)),
                        R2 = c(R2(lm1P, hideY, formula = 'traditional'),
                               R2(lm2P, hideY, formula = 'traditional'),
                               R2(lm3P, hideY, formula = 'traditional')),
                        MAE = c(MAE(lm1P, hideY), MAE(lm2P, hideY),
                                MAE(lm3P, hideY)))
kable(compTable, digits = 3L, format.args = list(big.mark = ','),
      caption = "Model Performance on Test Set")
```

# Discussion & Conclusions

@@

```{r loadLibraries}
```
```{r loadData}
```
```{r dataPrep}
```
```{r numLogDensities}
```
```{r factorPlot1}
```
```{r factorPlot2}
```

<!-- Feature Engineering & Selection -->
```{r subsetExtraction}
```
```{r addFeatures}
```

<!-- Model Building -->
```{r holdOut}
```
```{r lm1Train}
```
```{r lm1Table}
```
```{r lm2Train}
```
```{r lm2Table}
```
```{r lm3Train}
```

>>>>

# Methodology
Using the famous data set from Kaggle [@Mitchell2019TYV], we will explore
relationships between a video’s views and the number of likes, dislikes, and
comments using R [@RCT2020RAL]. We may also use a video’s country of origin and
the category or type of video as predictors.

@@

# Experimentation & Results
## Data Exploration
Since we are not factoring in time, it is incorrect to use all the observations.
Therefore, we will extract the latest observation by video by country and use
this subset.

```{r dataPrep}
# Select the row numbers of the first entry of the latest trending date by title
# and by country. There are 34 duplicates. Use this as our restricted data set
# corresponding to the most recent view count. The inner set of brackets gets
# the row number (called V1) and the outer set is a simple subset by those row
# numbers.
usExtract <- us_set[
  us_set[, .I[which.max(views)], by = c('title')]$V1
  ]
nobsUSExt <- dim(usExtract)[[1]]

# Substitute category for numeric ID through joining
usExtract <- us_cats[usExtract, on = 'id == category_id']

# Change data types for convenience, count tags and then get rid of actual tags
# and replace category IDs with actual categories
usExtract[, `:=`(trending_date = as.IDate(trending_date, format = "%y.%d.%m"),
                 views = as.double(views),
                 tag_count = 1L + str_count(tags, '[|]'),
                 id = NULL,
                 category = as.factor(category))
          ][, `:=`(tags = NULL,
                   category = relevel(category, 'Entertainment'))]

# Make name shorter for display purposes
setnames(usExtract,
         c('comment_count', 'tag_count', 'comments_disabled',
           'ratings_disabled', 'video_error_or_removed'),
         c('comments', 'tags', 'cmtDisabled', 'rtgDisabled', 'vidError'))

# Categorize the variable names
numVars <- c('views', 'likes', 'dislikes', 'comments', 'tags')

# Melt for numerics
usSetN <- melt(usExtract, measure.vars = numVars, variable.factor = FALSE,
               variable.name = 'measure', value.name = 'value')

# Melt for category
usSetF <- melt(usExtract, measure.vars = 'category', variable.factor = FALSE,
               variable.name = 'measure', value.name = 'value')

@@

### Categories
While tabular representation of factor predictors is difficult, a distribution
of videos by categories may prove informative. It is clear from the graphs below
that interests vary by category. `Entertainment` seems to be the most common in
the US with `Music` coming in second.

```{r catPlot, fig.height=4.5, fig.width=8}
ggplot(usSetF[measure == 'category'], aes(y = value)) +
  geom_bar() +
  ggtitle("Distribution by Category") +
  scale_x_continuous(labels = comma)
```

## Feature Selection & Engineering
We will consider the relationship between `views` and the numerical predictors
of `likes`, `dislikes`, `comments`, and `tags`. We will also consider the
`category`, whether or not comments or ratings were disabled and, if there was
an error with the video.

```{r likeDislikeCheck}
ldOutlier <- usExtract[!rtgDisabled & dislikes == 0 & likes == 0]
ldOn <- dim(ldOutlier)[[1]]
```

We will add two features. The first is `disp`: the disproportion between the
likes and dislikes. To minimize division by 0 errors, it is defined as follows:
\[
disp =\begin{cases}
\mathrm{ratings\;enabled}\quad\frac{\min(likes, dislikes)}{\max(likes, dislikes)}\\
\mathrm{ratings\;disabled}\quad 1
\end{cases}
\]
This constrains the ratio to \([0 \leq disp \leq 1]\) with a maximum when the
two are equal, and division by 0 only when both are 0. This usually occurs when
ratings are disabled. There is only `r ldOn` case out of the `r nobsUSExt`
observations where the ratings were not disabled, yet there are neither likes
nor dislikes. As this is a distinct incongruity for YouTube, we will remove that
one observation from the data. When ratings are disabled, perforce there is no
disparity so the ratio will be set to 1.

The second is `exceedMean`. This will be a factor if the views for the video
are greater than the *mean* views for all of the extracted data set. This will
also indicate the skewness of the data. Were it symmetrical, and we know it is
not from the graphs above, the mean should be near the median and this factor
should be close to equal for both 0 and 1.

```{r addFeatures}
# Remove the outlier
usExtract <- usExtract[!(!rtgDisabled & dislikes == 0 & likes == 0)]

# Precaclulate the mean views
usMean <- mean(usExtract$views)

# Add the engineered feature
usExtract[, `:=`(disp = ifelse(rtgDisabled, 1,
                               pmin(likes, dislikes) / pmax(likes, dislikes)),
                 exceedMean = factor(ifelse(views > usMean, 1, 0),
                                     levels = c(0, 1), labels = c(0, 1)))]
# Melt for exceedMean
usSetF <- melt(usExtract, measure.vars = 'exceedMean', variable.factor = FALSE,
               variable.name = 'measure', value.name = 'value')

# Plot the exceedMean variable
ggplot(usSetF[measure == 'exceedMean'], aes(x = value)) +
  geom_bar() +
  ggtitle("Proportion Above and Below Mean") +
  scale_y_continuous(labels = comma)
```

@@

@@

```{r holdout}
# Create seen and hidden sets
set.seed(617)
seenIDX <- createDataPartition(usExtract$views, p = 0.8)$Resample1
seenSet <- usExtract[seenIDX, ]
hideSet <- usExtract[-seenIDX, ]
```

### Simple Linear Regression
Linear regression may be the best known algorithm used when analyzing a
continuous numeric outcome. It searches for a linear relationship of the
predictors that minimizes the squared error between the "predictor" function and
the observations [@Sheather2009AMA].

This model will start with the numeric and logical features and the engineered
disparity ratio. It will not use the exceeding mean indicator, as that is
generated from the target variable, and it is felt it will distort the
prediction. It will then proceed through feature selection using the AIC as the
optimization metric.

```{r dummyVarsAA}
# Create dummy Variables. These will be used for the next three models
modDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
                   cmtDisabled + rtgDisabled + vidError + disp,
                   data = seenSet, fullRank = TRUE)
seenX <- predict(modDum, seenSet)
seenY <- seenSet$views
```

@@

### ElasticNet: Penalized Regression
Instead of using AIC to select features, one can make use of penalized
regression. Using an \(L_1\) penalty is the underpinnings of Lasso regression,
which can perform feature selection. Using a squared error term, \(L_2\), is at
the heart of ridge regression [@HTF2009TEo]. Using both methods together is
called the elastic net [@ZH2005RaV; ZH2020eEN]. To tune the hyperparameters,
which includes the weighting between Lasso and Ridge, we will use 10-fold
cross-validation.

```{r lm3Train}
trC <- trainControl(method = 'cv', number = 10L)
tG <- expand.grid(fraction = seq(0.85, 1, 0.001),
                  lambda = seq(0.01, 0.025, 0.001))
set.seed(181)
lm3 <- train(x = seenX, y = seenY,
             method = 'enet', trControl = trC, tuneGrid = tG)
```

@@

```{r modTest}
tstDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
                   cmtDisabled + rtgDisabled + vidError + disp,
                   data = hideSet, fullRank = TRUE)
hideX <- predict(tstDum, hideSet)
hideY <- hideSet$views
lm1P <- predict(lm1, newdata = hideX)
lm2P <- predict(lm2, newdata = hideX)
lm3P <- predict(lm3, newdata = hideX)
compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'ElasticNet'),
                        RMSE = c(RMSE(lm1P, hideY), RMSE(lm2P, hideY),
                                 RMSE(lm3P, hideY)),
                        R2 = c(R2(lm1P, hideY, formula = 'traditional'),
                               R2(lm2P, hideY, formula = 'traditional'),
                               R2(lm3P, hideY, formula = 'traditional')),
                        MAE = c(MAE(lm1P, hideY), MAE(lm2P, hideY),
                                MAE(lm3P, hideY)))
kable(compTable, digits = 3L, format.args = list(big.mark = ','),
      caption = "Model Performance on Test Set")
```

# Discussion & Conclusions
The three models contain both intuitive and counter-intuitive results. As
expected, videos with more ratings tend to have more views. However, as likes
outnumber dislikes, that tends to reduce the number of views. Having a video
witk

@@

```{r loadLibraries}
```
```{r loadData}
```
<!-- Data Exploration -->
```{r dataPrep}
```
```{r numLogDensities}
```
```{r catPlot}
```
```{r subsetExtraction}
```
```{r likeDislikeCheck}
```
```{r addFeatures}
```

<!-- Model Building -->
```{r holdout}
```
```{r dummyVarsAA}
```
```{r lm1Train}
```
```{r lm1Table}
```
```{r lm2Train}
```
```{r lm2Table}
```
```{r lm3Train}
```
